Accurate Measurement of Eight Bicycles' Physical Parameters

### 2010-06-01 report/parameterReport.tex
typed more on the report

\documentclass{article}
\usepackage[colorlinks=true, linkcolor=blue]{hyperref}
\usepackage{graphicx}

@@

\section{Introduction}
D\"{o}hring measured the physical parameters of a scooter~\cite{Dohring1953}.
Singh and Goel measured the physical parameters of a scooter~\cite{Singh1971}.
Roland and Massing measured the physical parameters of a bicycle in much the
same was as presented, including calculations of uncertainty from the indirect
measurement techniques~\cite{Roland1971}. This work is based off of the work
done by Kooijman~\cite{Kooijman2006}. I worked with Jodi Kooijman for a year and
was able to use much of the same apparatus and refine the measurement
technique.
\section{Accuracy}
Error propagation theory was utilized to calculate what the accuracy the
benchmark ideal parameters are based on the measurements taken.

@@

\subsection{Wheel radius}
The radius of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were
measured by measuring the linear distance traversed along the ground through
either 13 or 14 rotations of the wheel. Each wheel was measured separately and
the measurements were taken with a 72kg rider seated on the bicycle. A 30 meter
tape measure (resolution: 2mm) was taped on a flat level smooth floor. The tire
was marked with chalk and aligned with the tape measure. The accuracy of the
distance measurement is approximately $\pm0.005$m. The tires were pumped to the
recommended inflation pressure before the measurements. The wheel radius is
calculated by
\begin{equation}
	r=\frac{\textrm{distance}}{2\cdot\pi\cdot\textrm{rotations}}
	\label{eq:wheelRadius}
\end{equation}
\begin{figure}[tb]
	\begin{center}
		\includegraphics[width=4in]{images/tireChalk.jpg}
	\end{center}
	\caption{Wheel and tire with chalk mark aligned to the tape measure.}
	\label{fig:tireChalk}
\end{figure}
\begin{table}
	\begin{tabular}{llllllll}
	Bicycle   & B      & B*     & C      & G      & P      & S      & Y \& Y*\\
	Front\\
	Pressure  & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 4      \\
	Distance  & 28.060 & 27.980 & 27.980 & 29.044 & 29.366 & 27.772 & 27.925 \\
	Rotations & 13     & 13     & 13     & 14     & 14     & 13     & 13     \\
	Radius    & 0.344  & 0.343  & 0.343  & 0.330  & 0.334  & 0.340  & 0.342  \\
	Rear\\
	Pressure  & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 3.5    \\
	Distance  & 27.850 & 27.835 & 27.768 & 29.788 & 29.212 & 29.774 & 27.884 \\
	Rotations & 13     & 13     & 13     & 14     & 14     & 14     & 13     \\
	Radius    & 0.341  & 0.341  & 0.340  & 0.339  & 0.332  & 0.338  & 0.341	
	\end{tabular}
	\caption{}
	\label{tab:wheelRadius}
\end{table}
\subsection{Head tube angle}
The head tube angle was measured directly using an electronic level with a
$\pm0.2^{\circ}$ accuracy. The bicycle was fixed vertically, the front wheel
was aligned with the rear frame and the bicycle was unloaded. The steer axis
tilt $\lambda$ is the complement to the head tube angle.
\begin{equation}
	\lambda=\frac{\pi}{180}(90^{\circ}-\lambda_{ht})
\label{eq:headTubeAngle}
\end{equation}
\begin{table}
	\begin{tabular}{lllllllll}
	Bicycle   & B      & B*     & C      & G      & P      & S      & Y      & Y*\\
	Head tube angle & 67.1 & 67.1 & 69.0 & 71.1 & 74.2 & 73.1 & 72.7 & 70.6\\
	Steer axis tilt & 0.400 & 0.400 & 0.367 & 0.330 & 0.276 & 0.295 & 0.302 & 0.339
	\end{tabular}
	\caption{Head tube angle and steer axis tilt}
	\label{tab:lambda}
\end{table}

\subsection{Trail}
Trail is difficult to measure directly so we instead chose to measure the fork
offset. The fork offset was measured by clamping the steer tube of the front
fork into a v-block on a flat table. A ruler was used to measure the height of
the center of the head tube and the height of the center of the axle axis. The
fork blades where aligned such that the axle axis was parallel to the table
surface.
\begin{equation}
	c=\frac{r_\mathrm{F}\sin{\lambda}-f_o}{\cos{\lambda}}
	\label{eq:trail}
\end{equation}
\begin{table}
	\begin{tabular}{lllllllll}
	Bicycle     & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
	Fork offset & 0.071 & 0.071 & 0.045 & 0.039 & 0.032 & 0.045 & 0.057 & -0.057\\
	Trail       & 0.069 & 0.068 & 0.083 & 0.072 & 0.062 & 0.056 & 0.047 & 0.180
	\end{tabular}
	\caption{Head tube angle and steer axis tilt}
	\label{tab:trail}
\end{table}
\subsection{Wheelbase}
The wheelbase was measured directly using a tape measure while the bicycle was
in the same upright fixed position used when measuring the head tube angle.
\begin{table}
	\begin{tabular}{lllllllll}
	Bicycle    & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
	Wheel base & 1.121 & 1.121 & 1.101 & 1.070 & 0.989 & 1.037 & 1.089 & 0.985
	\end{tabular}
	\caption{Wheelbase values.}
	\label{tab:wheelbase}
\end{table}

@@

	\begin{tabular}{lllllllll}
		Bicycle     & B     & B*    & C     & G     & P    & S     & Y     & Y*\\
		total       & 18.5  & 23.5  & 20.9  & 10.5  & 9.9  & 17.6  & 10.2  & 10.2\\
		frame       & 9.86  & 14.71 & 9.18  & 4.48  & 4.49 & 7.22  & 3.31  & 3.31\\
		fork        &       & 3.22  & 4.57  & 2.52  & 2.27 & 3.04  & 2.45  & 2.45\\
		rear wheel  & 3.11  & 3.11  & 3.96  & 1.94  & 1.38 & 3.96  & 2.57  & 2.57\\
		front wheel & 2.02  & 2.02  & 3.55  & 1.50  & 1.58 & 3.33  & 1.90  & 1.90\\
		total sum   & 18.21 & 23.06 & 21.26 & 10.44 & 9.72 & 17.55 & 10.23 & 10.23

@@

compound and torsional pendulums and measuring the period of oscillation when

@@

>>>>

\documentclass{article}
\usepackage[lmargin=1in,rmargin=1in]{geometry}
\usepackage[colorlinks=true, linkcolor=blue]{hyperref}
\usepackage{graphicx}

@@

\section{Introduction}
D\"{o}hring measured the physical parameters of a scooter~\cite{Dohring1953}.
Singh and Goel measured the physical parameters of a scooter~\cite{Singh1971}.
Roland and Massing measured the physical parameters of a bicycle in much the
same was as presented, including calculations of uncertainty from the indirect
measurement techniques~\cite{Roland1971}. This work is based off of the work
done by Kooijman~\cite{Kooijman2006}. I worked with Jodi Kooijman for a year and
was able to use much of the same apparatus and refine the measurement
technique.
\section{Accuracy}
Error propagation theory was utilized to calculate what the accuracy the
benchmark ideal parameters are based on the measurements taken. If $x$ is a
random variable and is defined as $x=f(u,v,\ldots)$ then the sample variance of
$x$ is defined as
\begin{equation}
    s_x^2 = \frac{1}{N-1}\sum^N_{i=1}
    \left[(u_i - \bar{u})^2\left(\frac{\partial x}{\partial u}\right)^2 +
    (v_i - \bar{v})^2\left(\frac{\partial x}{\partial v}\right)^2 +
    2(u_i - \bar{u})(v_i - \bar{v})\left(\frac{\partial x}{\partial u}\right)\left(\frac{\partial x}{\partial v}\right)
    + \ldots\right]
    \label{eqn:sampleVariance}
\end{equation}
Using the definitions for variance and covariance
Equation~\ref{eqn:sampleVariance} can be simplified to
\begin{equation}
    s_x^2 = s_u^2\left(\frac{\partial x}{\partial u}\right)^2 +
            s_v^2\left(\frac{\partial x}{\partial v}\right)^2 +
            2s_{uv}\left(\frac{\partial x}{\partial u}\right)\left(\frac{\partial x}{\partial v}\right)
            + \ldots
    \label{eqn:variance}
\end{equation}
If $u$ and $v$ are uncorrelated then $s_{uv}=0$. We assume that the all the
random variables hereafter are uncorrelated. Equation~\ref{eqn:variance} can be
used to calculated the variance of all types of functions. The most basic are
functions such as addtion
\begin{eqnarray}
    \label{eqn:addition}
    x &=&  au + bv\\
    s_x &=& a^2s_u^2 + b^2s_v^2
\end{eqnarray}

@@

\subsection{Wheel radius}
The radius of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were
measured by measuring the linear distance traversed along the ground through
either 13 or 14 rotations of the wheel. Each wheel was measured separately and
the measurements were taken with a 72kg rider seated on the bicycle. A 30 meter
tape measure (resolution: 2mm) was taped on a flat level smooth floor. The tire
was marked with chalk and aligned with the tape measure. The accuracy of the
distance measurement is approximately $\pm0.005$m. The tires were pumped to the
recommended inflation pressure before the measurements. The wheel radius is
calculated by
\begin{equation}
	r=\frac{\textrm{distance}}{2\cdot\pi\cdot\textrm{rotations}}
	\label{eq:wheelRadius}
\end{equation}
\begin{figure}[tb]
	\begin{center}
		\includegraphics[width=4in]{images/tireChalk.jpg}
	\end{center}
	\caption{Wheel and tire with chalk mark aligned to the tape measure.}
	\label{fig:tireChalk}
\end{figure}
\begin{table}
	\begin{tabular}{llllllll}
	Bicycle   & B      & B*     & C      & G      & P      & S      & Y \& Y*\\
    \hline
	Front\\
	Pressure  & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 4      \\
	Distance  & 28.060 & 27.980 & 27.980 & 29.044 & 29.366 & 27.772 & 27.925 \\
	Rotations & 13     & 13     & 13     & 14     & 14     & 13     & 13     \\
	Radius    & 0.344  & 0.343  & 0.343  & 0.330  & 0.334  & 0.340  & 0.342  \\
	Rear\\
	Pressure  & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 3.5    \\
	Distance  & 27.850 & 27.835 & 27.768 & 29.788 & 29.212 & 29.774 & 27.884 \\
	Rotations & 13     & 13     & 13     & 14     & 14     & 14     & 13     \\
	Radius    & 0.341  & 0.341  & 0.340  & 0.339  & 0.332  & 0.338  & 0.341
	\end{tabular}
	\caption{}
	\label{tab:wheelRadius}
\end{table}
\subsection{Head tube angle}
The head tube angle was measured directly using an electronic level with a
$\pm0.2^{\circ}$ accuracy. The bicycle was fixed vertically, the front wheel
was aligned with the rear frame and the bicycle was unloaded. The steer axis
tilt $\lambda$ is the complement to the head tube angle.
\begin{equation}
	\lambda=\frac{\pi}{180}(90^{\circ}-\lambda_{ht})
\label{eq:headTubeAngle}
\end{equation}
\begin{table}
	\begin{tabular}{lllllllll}
    \hline
	Bicycle   & B      & B*     & C      & G      & P      & S      & Y      & Y*\\
	Head tube angle & 67.1 & 67.1 & 69.0 & 71.1 & 74.2 & 73.1 & 72.7 & 70.6\\
	Steer axis tilt & 0.400 & 0.400 & 0.367 & 0.330 & 0.276 & 0.295 & 0.302 & 0.339
	\end{tabular}
	\caption{Head tube angle and steer axis tilt}
	\label{tab:lambda}
\end{table}

\subsection{Trail}
Trail is difficult to measure directly so we instead chose to measure the fork
offset. The fork offset was measured by clamping the steer tube of the front
fork into a v-block on a flat table. A ruler was used to measure the height of
the center of the head tube and the height of the center of the axle axis. The
fork blades where aligned such that the axle axis was parallel to the table
surface.
\begin{equation}
	c=\frac{r_\mathrm{F}\sin{\lambda}-f_o}{\cos{\lambda}}
	\label{eq:trail}
\end{equation}
\begin{table}
	\begin{tabular}{lllllllll}
	Bicycle     & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
    \hline
	Fork offset & 0.071 & 0.071 & 0.045 & 0.039 & 0.032 & 0.045 & 0.057 & -0.057\\
	Trail       & 0.069 & 0.068 & 0.083 & 0.072 & 0.062 & 0.056 & 0.047 & 0.180
	\end{tabular}
	\caption{Head tube angle and steer axis tilt}
	\label{tab:trail}
\end{table}
\subsection{Wheelbase}
The wheelbase was measured directly using a tape measure while the bicycle was
in the same upright fixed position used when measuring the head tube angle.
\begin{table}
	\begin{tabular}{lllllllll}
	Bicycle    & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
    \hline
	Wheel base & 1.121 & 1.121 & 1.101 & 1.070 & 0.989 & 1.037 & 1.089 & 0.985
	\end{tabular}
	\caption{Wheelbase values.}
	\label{tab:wheelbase}
\end{table}

@@

    \centering
	\begin{tabular}{llllllllll}
		Bicycle &     $\sigma_m$ & B     & B*    & C     & G     & P    & S     & Y     & Y*\\
        \hline
		total   &    0.1       & 18.5  & 23.5  & 20.9  & 10.5  & 9.9  & 17.6  & 10.2  & 10.2\\
        \hline
		frame   &    0.02      & 9.86  & 14.71 & 9.18  & 4.48  & 4.49 & 7.22  & 3.31  & 3.31\\
		fork    &    0.02      & 3.22  & 3.22  & 4.57  & 2.52  & 2.27 & 3.04  & 2.45  & 2.45\\
		rear wheel & 0.02      & 3.11  & 3.11  & 3.96  & 1.94  & 1.38 & 3.96  & 2.57  & 2.57\\
		front wheel& 0.02      & 2.02  & 2.02  & 3.55  & 1.50  & 1.58 & 3.33  & 1.90  & 1.90\\
		total sum  & 0.04      & 18.21 & 23.06 & 21.26 & 10.44 & 9.72 & 17.55 & 10.23 & 10.23

@@

compound and torsional pendulums and measuring their period of oscillation when

@@

\begin{figure}[tbp]
        \includegraphics[width=5in]{../data/pendDat/graphs/BrowserFrameCompoundFirst1.png}
    \caption{Example of the raw voltage data taken during a 30 second
    measurement of the oscillation of one of the components.}
    \label{fig:voltage}
